---
title: 'Reproducible Research: Week2'
author: "Sweta Mohanty"
date: "10/04/2018"
output: html_document
editor_options: 
  chunk_output_type: inline
---
#### Reading the data
```{r, include=TRUE,echo=TRUE}
analysis<-read.csv("activity.csv",header=TRUE)
```
#### Transforming the data
```{r,include=TRUE,echo=TRUE }
analysis$date<-as.Date(as.character(analysis$date))
analysisNA<-is.na(analysis$steps)
newanalysis<-analysis[!analysisNA,]
```
### What is mean total number of steps taken per day?
#### Total no of steps per day

```{r,include=TRUE,echo=TRUE }
totalsteps<-aggregate(newanalysis$steps,by=list(newanalysis$date),sum)
names(totalsteps)[1]<-"Date"
names(totalsteps)[2]<-"Steps"
totalsteps
```

#### Plotting the histogram
```{r,include=TRUE,echo=TRUE }
hist(totalsteps$Steps,breaks= nrow(totalsteps),main="Histogram for Total Steps",xlab="Total Steps",col="Green")
```

#### Mean of Total Steps
```{r,include=TRUE,echo=TRUE }
mean(totalsteps$Steps)
```
#### Median of Total Steps
```{r,include=TRUE,echo=TRUE}
median(totalsteps$Steps)
```
### What is the average daily activity pattern?
#### Average steps per interval

```{r,include=TRUE,echo=FALSE}
dailyactivity<-aggregate(newanalysis$steps,by=list(newanalysis$interval),sum)
names(dailyactivity)[1]<-"Interval"
names(dailyactivity)[2]<-"Steps"
```

#### Plotting the activity

```{r,include=TRUE,echo=TRUE}
plot(dailyactivity$Interval,dailyactivity$Steps,type="l",main="Daily Activity Pattern",xlab="5 min interval",ylab="Average daily steps")
```

#### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r,include=TRUE,echo=TRUE}
maxinterval<-dailyactivity[which.max(dailyactivity$Steps),]
maxinterval
```
### Imputing missing values

#### Total no. of missing values
```{r,include=TRUE,echo=TRUE}
missingrow<-sum(analysisNA)
missingrow
```

#### Replacing the missing values with median steps of the time interval

```{r,include=TRUE,echo=TRUE}
analysis1<-analysis
medianval<-tapply(newanalysis$steps,newanalysis$interval,mean,na.rm=TRUE,simplify = TRUE)
analysis1$steps[analysisNA]<-medianval[as.character(analysis1$interval[analysisNA])]
sum(is.na(analysis1$steps))
```
#### Plotting the activity with filled in missing steps

```{r,include=TRUE,echo=TRUE}
totalsteps1<-aggregate(analysis1$steps,by=list(analysis1$date),sum)
names(totalsteps1)[1]<-"Date"
names(totalsteps1)[2]<-"Steps"
hist(totalsteps1$Steps,breaks= nrow(totalsteps1),main="Histogram for Total Steps",xlab="Total Steps",col="Green")
```

#### Mean of Total steps
```{r,include=TRUE,echo=TRUE}
mean(totalsteps1$Steps)
```

#### Median of Total steps

```{r,include=TRUE,echo=TRUE}
median(totalsteps1$Steps)
```
#### The mean of the daily total steps are same for both data sets. But the median has changed once the missings values were filled in.

### Are there differences in activity patterns between weekdays and weekends?
```{r,include=TRUE,echo=TRUE}
analysis1$Daytype<-weekdays(analysis1$date)
analysis1$Daytype<-ifelse(analysis1$Daytype=="Saturday"|analysis1$Daytype=="Sunday","Weekend","Weekday")
```
#### Mean of Steps by  Interval and day type

```{r,include=TRUE,echo=TRUE,eval=TRUE}
library(ggplot2)
meandaytype<-aggregate(analysis1$steps,by=list(analysis1$Daytype,analysis1$interval),mean)
names(meandaytype)[1]<-"Daytype"
names(meandaytype)[2]<-"Interval"
names(meandaytype)[3]<-"steps"

p<-ggplot(meandaytype,aes(x=Interval,y= steps,color=Daytype)) + geom_line()+facet_grid(Daytype~.)+labs(title="Mean steps per inteval and day type",x="Interval",y="MeanSteps")
print(p)
```

